Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-20 paper/paper.tex
No need to special-case loops

We introduce definitions and algorithms on undirected and unlabelled graphs without loops; the
extension to general graphs is straightforward and is discussed in \cref{extension}. An undirected
graph $G$ is defined by a finite set of vertices $\operatorname{V}(G)$ and a set of undirected edges
$\operatorname{E}(G) \subseteq \operatorname{V}(G) \times \operatorname{V}(G)$, where $(u, v) \in
\operatorname{E}(G) \Rightarrow (v, u) \in \operatorname{E}(G) \wedge u \ne v$. The neighbourhood of
a vertex $v$, written $\operatorname{N}(G, v)$, is the set of vertices to which it is adjacent.
Given a graph $G$, two vertices $v_s, v_e \in \operatorname{V}(G)$ are connected by a path in $G$ if
there exists a sequence of vertices $(v_0, v_1, \ldots, v_k)$ such that $v_0 = v_s$, $v_k = v_e$,
and each pair of successive vertices is connected by an edge, i.e.\ $\forall i\in \{1 \ldots k \},
(v_{i-1}, v_i) \in \operatorname{E}(G)$. A graph $G$ is \emph{connected} if every pair of vertices
is connected by a path.

@@

There are two main approaches for solving MCS. The first approach (described in \cref{CP}) is based
on CP, whilst the second (described in \cref{clique}) is based on a reformulation of MCS into a
maximum clique problem. Both approaches are described for undirected, unlabelled graphs without
loops; their extension to richer graphs is discussed in \cref{extension}.  (Other approaches have
been tried, mixed integer programming \cite{DBLP:journals/anor/PivaS12} and heuristics
\cite{DBLP:journals/jcisd/EnglertK15}; SAT encodings seem to struggle even for subgraph isomorphism
\cite{UpcomingIJCAIPaper}.)

@@

\subsection{Extension to Labelled or Directed Graphs, and Loops}\label{extension}

@@

The extension of MCS algorithms to directed graphs, where isomorphisms must preserve directed edges,
is similarly straightforward.
Finally, some datasets contain loops (self-adjacent vertices). In this case vertices with loops must
be mapped to vertices with loops (with the same labels, if appropriate), and vertices without loops
must be mapped to vertices without loops.

Labels, directed edges and loops usually simplify the solution process, both for CP and clique-based
approaches: vertex labels reduce domain sizes for CP, and the number of matching nodes in the
association graphs; edge labels tighten edge constraints for CP, and make the association graph
sparser for clique-based approaches. It is worth noting that edge constraints do not help CP
approaches to do more filtering so long as $\bot$ remains in variable domains: every pair of
variables $(x_i, x_j)$ having  $D(x_i) \neq \emptyset$ and $\bot \in D(x_j)$ is arc consistent,
since $\bot$ is a support for any value $u \in D(x_i)$. However, as soon as $\bot$ is removed from
domains (i.e.\ when the number of variables assigned to $\bot$ has reached the best known bound on
the size of the MCS), maintaining arc consistency may filter values, and then the tighter constraints
increase the opportunities for filtering.

>>>>

We introduce definitions and algorithms on undirected and unlabelled graphs without loops; the
extension to general graphs is straightforward and is discussed in \cref{extension}. An undirected
graph $G$ is defined by a finite set of vertices $\operatorname{V}(G)$ and a set of undirected edges
$\operatorname{E}(G) \subseteq \operatorname{V}(G) \times \operatorname{V}(G)$, where $(u, v) \in
\operatorname{E}(G) \Rightarrow (v, u) \in \operatorname{E}(G)$. The neighbourhood of
a vertex $v$, written $\operatorname{N}(G, v)$, is the set of vertices to which it is adjacent.
Given a graph $G$, two vertices $v_s, v_e \in \operatorname{V}(G)$ are connected by a path in $G$ if
there exists a sequence of vertices $(v_0, v_1, \ldots, v_k)$ such that $v_0 = v_s$, $v_k = v_e$,
and each pair of successive vertices is connected by an edge, i.e.\ $\forall i\in \{1 \ldots k \},
(v_{i-1}, v_i) \in \operatorname{E}(G)$. A graph $G$ is \emph{connected} if every non-adjacent pair
of vertices is connected by a path.

@@

There are two main approaches for solving MCS. The first approach (described in \cref{CP}) is based
on CP, whilst the second (described in \cref{clique}) is based on a reformulation of MCS into a
maximum clique problem. Both approaches are described for undirected, unlabelled graphs; their
extension to richer graphs is discussed in \cref{extension}.  (Other approaches have been tried,
mixed integer programming \cite{DBLP:journals/anor/PivaS12} and heuristics
\cite{DBLP:journals/jcisd/EnglertK15}; SAT encodings seem to struggle even for subgraph isomorphism
\cite{UpcomingIJCAIPaper}.)

@@

\subsection{Extension to Labelled or Directed Graphs}\label{extension}

@@

The extension of MCS algorithms to directed graphs, where isomorphisms must preserve directed edges,
is similarly straightforward.

Labels and directed edges usually simplify the solution process, both for CP and clique-based
approaches: vertex labels reduce domain sizes for CP, and the number of matching nodes in the
association graphs; edge labels tighten edge constraints for CP, and make the association graph
sparser for clique-based approaches. It is worth noting that edge constraints do not help CP
approaches to do more filtering so long as $\bot$ remains in variable domains: every pair of
variables $(x_i, x_j)$ having  $D(x_i) \neq \emptyset$ and $\bot \in D(x_j)$ is arc consistent,
since $\bot$ is a support for any value $u \in D(x_i)$. However, as soon as $\bot$ is removed from
domains (i.e.\ when the number of variables assigned to $\bot$ has reached the best known bound on
the size of the MCS), maintaining arc consistency may filter values, and then the tighter constraints
increase the opportunities for filtering.
